### 2021-03-31 code/import_pew_data.Rmd
cleaning

---
title: "Pew Cleaning"
author: "Jake Chanenson"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(foreign)      
library(ggplot2)      
library(colorspace)   #better colors for ggplot
library(tidyverse)    
library(dplyr)
library(gridExtra)
library(viridis)
library(ggridges)
library(socviz)
library(broom)
library(stargazer)
library(corrr)
library(GGally)
library(haven) #needed to import stada data
```

## Initial Data Import
```{r, include=FALSE}
if(!file.exists("../../Pew/pewTech.RData") ) { #skip import if RData file exists
# Data set from Pew 
importedData <- read_sav("../../Pew/ATP W49.sav")

# Add the QKEY col
pew <- importedData$QKEY

# Iterate through data set
for(i in 2:(ncol(importedData)-1)){ #skip the QKEY column (1st col) and the WEIGHT_W49 col (last col)
        temp <- as_factor(importedData[ , i])
        pew <- cbind(pew,temp)
        temp <- 0
}

# Add the WEIGHT_W49 col
temp <- as_factor(importedData[ , ncol(importedData)])
pew <- cbind(pew,temp)

# Show pew
head(pew)

# Save the file

    write.csv(pew, "../../Pew/Pew_Tech.csv")
    save(pew, file = "../../Pew/pewTech.RData")
}

load("../../Pew/pewTech.RData")
```

## Sample Viz
```{r, echo=FALSE}
load("../../Pew/pewTech.RData")

pew %>% drop_na(TRACKCO1a_W49) %>% ggplot(aes(x = TRACKCO1a_W49, fill = HOMEASSIST1_W49)) +
        geom_bar(position = "dodge", color = 'black') +
        theme(axis.text.x = element_text(angle=25, hjust = 1)) +
        labs( title = 'Dodged Bar Chart: \nDo Smart Speaker Owners Think They Are Being Tracked Online',
            x = "Opnion on How Much of What You Do Online is Being Tracked",
            y = "Count",
            fill = "Smart Speaker Ownership",
            caption = "Pew Research Data")
ggsave(file="../imgs/hi.png", width = 12, height = 10, units = "in")
```

```{r, echo=FALSE}
pew %>% drop_na(TRACKCO1a_W49) %>%  filter(HOMEASSIST1_W49 != "Refused") %>% ggplot(aes(x = TRACKCO1a_W49, fill = TRACKCO1a_W49)) +
        geom_bar(color = 'black') +
        facet_wrap(HOMEASSIST1_W49~., nrow = 2) +
        geom_text(stat='count', aes(label=..count..), vjust= -.5) +
        ylim(0,800) +
        theme(axis.text.x = element_text(angle=25, hjust = 1)) +
        labs( title = 'Faceted Bar Chart: \nDo Smart Speaker Owners Think They Are Being Tracked Online',
            x = "Opnion on How Much of What You Do Online is Being Tracked",
            y = "Count",
            fill = "Smart Speaker Ownership",
            caption = "Pew Research Data")

ggsave(file="../imgs/faceted_pew.png", width = 12, height = 10, units = "in")
```