### 2021-04-21 code/Viz Creation.Rmd
Update Viz Creation.Rmd

---
title: "Viz Creation"
author: "Jake Chanenson"
date: "4/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(foreign)      
library(ggplot2)      
library(colorspace)   #better colors for ggplot
library(tidyverse)    
library(dplyr)
library(gridExtra)
library(viridis)
library(ggridges)
library(socviz)
library(broom)
library(stargazer)
library(corrr)
library(GGally)
library(haven) #needed to import stada data
library(RColorBrewer)
```

```{r load-data}
load("../../Pew/pewTech.RData")
```

## `Know` Dummy Var
Generate counts and props of how many people get the following questions right

* KNOW9

* KNOW8

* KNOW7

* KNOW4

* KNOW3

* KNOW1

```{r}
## Generate count dummies
count_df <- pew %>% mutate(K1T = ifelse(KNOW1_W49 == "Can track your visits and activity on the site", yes = 1, no = 0 ),
               K3T = ifelse(KNOW3_W49 == "Has created a contract between itself and its users about how it will use their data", yes = 1, no = 0),
               K4T = ifelse(KNOW4_W49 == "Information entered into the site is encrypted", yes = 1, no = 0),
               K7T = ifelse(KNOW7_W49 == "Internet service providers should treat all traffic on their networks equally", yes = 1, no = 0),
               K8T = ifelse(KNOW8_W49 == "A coworker who uses the same computer", yes = 1, no = 0),
               K9T = ifelse(KNOW9_W49 == "Picture of 2-step code", yes = 1, no = 0)
               ) %>% select(K1T, K3T, K4T, K7T, K8T, K9T)
count_df

## Sum count dummies, add to pew data set
pew <- pew %>% mutate(sumKnow134789 = count_df$K1T+count_df$K3T+count_df$K4T+count_df$K7T+count_df$K8T+count_df$K9T,
               propKnow134789 = sumKnow134789/6,
               .after = TOTALKNOW_W49)
```

```{r}
pew %>% ggplot(aes(x = sumKnow134789)) +
  geom_bar()
```

```{r}
dum <- as_factor(pew$sumKnow134789)
pew <- pew %>% mutate(sumKnowFactor = dum, .after = sumKnow134789)

foo1 <- levels(pew$sumKnowFactor) #why can't this be a quant var

pew %>% drop_na(DATAUSEd_W49) %>% ggplot(aes(x = DATAUSEd_W49, fill = sumKnowFactor)) +
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Dark2", limits = rev(foo1))
```

```{r}
pew %>% drop_na(DATAUSEd_W49) %>% ggplot(aes(x = DATAUSEd_W49, y = propKnow134789)) +
  geom_boxplot(outlier.colour="red", outlier.shape=16,
             outlier.size=2, notch=FALSE) +
  geom_jitter(shape=1, position=position_jitter(0.2)) +
  scale_y_continuous(label = scales::percent) +
  labs (
    title = "Tech Literacy VS X ",
    x = " Fitness Trackers Data Sharing With Medical Researchers\nseeking to better understand the link between exercise and heart disease",
    y = "Prop"
  )
```

```{r, echo = echoVal, warning= F, message = F}

pew <- pew %>% mutate(HOMEASSIST1_num = ifelse(HOMEASSIST1_W49 == "Yes", yes = 1, 
                                         ifelse(HOMEASSIST1_W49 == "No", yes = 0, no  = NA)))
pew %>% drop_na(DATAUSEd_num) %>%  
  ggplot(aes(x= propKnow134789, y=DATAUSEd_num)) +
  geom_jitter(aes(color = DATAUSEd_num), alpha=.3, size =1, height = .1, width = .005)+
  geom_smooth(method = "glm",
    method.args = list(family = "binomial"), se = T, color ='black') +
  # scale_color_manual(values = c('blue','red'))+
  theme_bw()
```

```{r}

pew <- pew %>% mutate(DATAUSEd_num = ifelse(DATAUSEd_W49 == "Acceptable", yes = 1, 
                                         ifelse(DATAUSEd_W49 == "Unacceptable", yes = 0, 
                                                ifelse(DATAUSEd_W49 == "Not sure", yes = -1, no  = NA))),
                      DATAUSEd_fct = as_factor(DATAUSEd_W49))
pew %>% drop_na(DATAUSEd_num) %>% filter(DATAUSEd_num != -1) %>%  
  ggplot(aes(x= propKnow134789, y=DATAUSEd_num)) +
  geom_jitter(aes(color = DATAUSEd_fct), alpha=.3, size =1, height = .1, width = .005)+
  geom_smooth(method = "glm",
    method.args = list(family = "binomial"), se = T, color ='black') +
  # scale_color_manual(values = c('blue','red'))+
  theme_bw()

```