### 2021-05-07 code/rtbf.rmd

---
title: "RTBF Visualization"
output:
  pdf_document: default
word_document: default
html_document: default
---

### Author: Adriana Knight

```{r load-libraries, echo = F, results = 'hide', message=F, warning = F}
library(tinytex)
library(knitr)
library(ggplot2)   #plotting package
library(gridExtra)  #allow side by side plots
library(tidyr)   #a widely used, but relatively new, package for producing tidier data and tidier output
library(broom)        #a widely used, but relatively new, package for producing tidier data and tidier output; part of tidyverse
library(dplyr)
echoVal <- F   #change this to FALSE for presentations where you don't want to show your code in the knitted document
```

```{r dataSetup, echo = echoVal}
load("../../Pew/pewTech.RData")
```

```{r moreDataSetup, echo = echoVal}
## Generate RTBF count dummies 
count_rtbf <- pew %>% mutate(forget_a = ifelse(RTBFa_W49 == "Yes, should be able to remove this from online searches", yes = 1, no = 0),
                             forget_b = ifelse(RTBFb_W49 == "Yes, should be able to remove this from online searches", yes = 1, no = 0),
                             forget_c = ifelse(RTBFc_W49 == "Yes, should be able to remove this from online searches", yes = 1, no = 0),
                             forget_d = ifelse(RTBFd_W49 == "Yes, should be able to remove this from online searches", yes = 1, no = 0)
                             ) %>% select(forget_a, forget_b, forget_c, forget_d)

## Sum RTBF count dummies, add to pew data set
pew <- pew %>% mutate(sumRTBF = count_rtbf$forget_a + count_rtbf$forget_b + count_rtbf$forget_c + count_rtbf$forget_d,
                      propRTBF = sumRTBF/4,
                      .after = RTBFd_W49)
```

```{r central_fig, echo = echoVal}
levels(pew$PUBLICDATA_W49) <- c("Find others' information","Keep their own information private", "Refused")
pew %>% drop_na(PUBLICDATA_W49) %>% drop_na(sumRTBF) %>% filter(PUBLICDATA_W49 != "Refused") %>% ggplot(aes(x = PUBLICDATA_W49, fill = as.factor(sumRTBF))) +
  geom_bar(position = "dodge", color = 'black') +
  labs (
    title = "Attitudes on How Many Categories of Data Should Be Protected by\n\"The Right To Be Forgotten\" by Stance on Personal Data",
    x = "Participants find it more important to...",
    y = "",
    caption = "Data sourced from\n2019 PEW American Trends Panel\nSurvey on Social Media Usage and\nPerceptions of Digital Privacy"
  ) +
  scale_fill_brewer(palette="YlGnBu", name = "How many of the following\ncategories of data should\nfall under the Right To Be Forgotten:\n   -Data collected by law enforcement\n   -Employment history or work record\n   -Negative media coverage\n   -Potentially embarassing photos or\n     videos", labels=c("None","One","Two","Three","Four"))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

```