### 2021-05-17 code/Misc_Viz.Rmd
:stars:

---
title: "Misc Viz"
author: "Jake Chanenson"
date: "5/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(foreign)      
library(ggplot2)      
library(colorspace)   #better colors for ggplot
library(tidyverse)    
library(dplyr)
library(grid)
library(gridExtra)
library(broom)
library(GGally)
library(RColorBrewer)
library(likert) #for creating a likert plot
library(kableExtra) # not needed but makes the table nicer
echoval <- F
load("../../Pew/pewTech.RData")
```

### Concern Of Company Tracking Faceted on Tech Literacy
```{r company tracking, echo= echoval, message = FALSE, warning = FALSE}
pew %>% drop_na(CONCERNCO_W49) %>% filter(CONCERNCO_W49 != "Refused") %>% ggplot(aes(x = CONCERNCO_W49, fill = CONCERNCO_W49)) +
  geom_bar(color = 'black') +
  theme(axis.text.x = element_blank()) +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(. ~groupKnow, scales = "free_y") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank()) +
  labs (
    title = "Concern Of Company Tracking Grouped by Tech Literacy",
    fill = "How Concerned Are Respondents\nAbout Companies Tracking Them?",
    x = "Tech Literacy Level",
    caption = "Data sourced from\n2019 PEW American Trends Panel\nSurvey on Social Media Usage and\nPerceptions of Digital Privacy"
  ) 
ggsave("../imgs/Misc/company_tracking.png")
```

### Understanding of Targeted Ad Data Grouped by Tech Literacy
```{r targeted-ad, echo= echoval, message = FALSE, warning = FALSE}
pew %>% drop_na(PROFILE4_W49) %>% filter(PROFILE4_W49 != "Refused") %>% ggplot(aes(x = PROFILE4_W49, fill = PROFILE4_W49)) +
  geom_bar(color = 'black') +
  facet_wrap(. ~groupKnow, scales = "free_y") +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank()) +
  labs (
    title = "Understanding of Targeted Ad Data Grouped by Tech Literacy",
    # subtitle = "Optional",
    fill = "Do You Understand What Data\nIs Used To Create Targeted Ads?",
    x = "Tech Literacy Level",
    caption = "Data sourced from\n2019 PEW American Trends Panel\nSurvey on Social Media Usage and\nPerceptions of Digital Privacy"
  )
ggsave("../imgs/Misc/targeted_ad_understanding.png")
```

### Key Figure 2 With Upper and Lower Percents
```{r, KF2-percents, echo= echoval, message = FALSE, warning = FALSE}
## Centerd, Percents 
## A bug in the likert package is what is alphabetizing the facets
plot(likert(df_f3.reverse[,c(1:3)], grouping = df_f3.reverse[,4]), legend.position="right", centered = TRUE, plot.percents=TRUE, plot.percent.low=TRUE, plot.percent.high=TRUE, plot.percent.neutral=FALSE, center = 2.5) + 
  scale_fill_manual(values = brewer.pal(n=4,"RdBu"), breaks = c("Not at all", "Not too much", "Somewhat", "A great deal")) + #order and color the likert boxes
  ggtitle(title2) +
  labs(caption = "Data sourced from\n2019 PEW American Trends Panel\nSurvey on Social Media Usage and\nPerceptions of Digital Privacy")+
  guides(fill = guide_legend(title="Responses", reverse = TRUE)) #to reverse the order in the legend
ggsave(file = "../imgs/Misc/KF2_percents.png")
```